Word Identification Under Multimodal Uncertainty

### 2018-02-17 Writeup/ms.Rmd
changes to journal write-up

affiliation:
  - id            : "1"
    institution   : "Department of Psychoogy, Stanford University"

@@

abstract: |
  Identifying the visual referent of a spoken word -- that a particular insect is referred to by the word ``bee'' -- requires both the ability to process and integrate multimodal input and the ability to reason under uncertainty. How do these tasks interact with one another? We introduce a task that allows us to examine how adults identify words under joint uncertainty in the auditory and visual modalities. We propose an ideal observer model of the task which provides an optimal account of how auditory and visual cues are combined. Model predictions are tested in three experiments where word recognition is made under two kinds of uncertainty: category ambiguity and/or distorting noise. In all cases, the optimal model explains much of the variance in human judgments. In particular, when the signal was not distorted with noise, participants weighted the auditory and visual cues optimally, that is, according to the relative reliability of each modality. But when one modality had noise added to it, human perceivers systematically preferred the unperturbed modality to a greater extent than the optimal model did. We discuss the broad implications of the results, especially in relation to early language acuquisition.

@@

@@

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library("papaja")
library(readr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(boot)
library(dplyr)
library(langcog)
theme_set(theme_bw())
```

@@

#Data from the 3 experiments 
#############################
exp1 <- read_delim("../Data_and_analysis/data_exp1_Anonym.txt", delim = " ") %>%
  filter(type == "Task") %>%
  mutate(experiment='Experiment1')

exp2 <- read_delim("../Data_and_analysis/data_exp2_Anonym.txt", delim = " ") %>%
  filter(type == "Task") %>%
  mutate(experiment='Experiment2')

exp3 <- read_delim("../Data_and_analysis/data_exp3_Anonym.txt", delim = " ") %>%
  filter(type == "Task") %>%
  mutate(experiment='Experiment3')
```

```{r}
#First deal with all data
```

@@

#Filter who took the experiment more than once

#Exp1
 token_data_1 <- exp1 %>%
   group_by(ID) %>%
   dplyr::summarise(N_tokens=n())

 N_twice_1 <- token_data_1 %>%
   filter(N_tokens > 70) %>%
   nrow()

 exp1_once <- exp1 %>%
   left_join(token_data_1) %>%
   filter(N_tokens == 70) #every subject is supposed to produce exactly 70 tokens

 #Exp2
token_data_2 <- exp2 %>%
   group_by(ID) %>%
   dplyr::summarise(N_tokens=n())

 N_twice_2 <- token_data_2 %>%
   filter(N_tokens > 70) %>%
   nrow()

 exp2_once <- exp2 %>%
   left_join(token_data_2) %>%
   filter(N_tokens == 70)

 #Exp3
token_data_3 <- exp3 %>%
   group_by(ID) %>%
   dplyr::summarise(N_tokens=n())

 N_twice_3 <- token_data_3 %>%
   filter(N_tokens > 70) %>%
   nrow()

 exp3_once <- exp3 %>%
   left_join(token_data_3) %>%
   filter(N_tokens == 70)

@@

#Exp1
exp1_noProb <- exp1_once %>%
  filter(problem=="No")

@@

#Exp2
exp2_noProb <- exp2_once %>%
  filter(problem=="No")

@@

#Exp3
exp3_noProb <- exp3_once %>%
  filter(problem=="No")

@@

#Summary
sounds_exp1 <- exp1_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  #multi_boot_standard(col = "answer")
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

concepts_exp1 <- exp1_good %>%
  filter(condition == "concept") %>%
  group_by(concept_dist) %>%
  #multi_boot_standard(col = "answer")
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Visual") %>%
  rename(distance = concept_dist)

joint_exp1 <- exp1_good %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  #multi_boot_standard(col = "answer")
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Joint")

@@

#Summary
sounds_exp2 <- exp2_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  #multi_boot_standard(col = "answer")
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 2", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

concepts_exp2 <- exp2_good %>%
  filter(condition == "concept") %>%
  group_by(concept_dist) %>%
  #multi_boot_standard(col = "answer")
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 2", 
         Condition="Visual") %>%
  rename(distance = concept_dist)

joint_exp2 <- exp2_good %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  #multi_boot_standard(col = "answer")
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 2", 
         Condition="Joint")

@@

#Summary
sounds_exp3 <- exp3_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  #multi_boot_standard(col = "answer")
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 3", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

concepts_exp3 <- exp3_good %>%
  filter(condition == "concept") %>%
  group_by(concept_dist) %>%
  #multi_boot_standard(col = "answer")
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 3", 
         Condition="Visual") %>%
  rename(distance = concept_dist)

joint_exp3 <- exp3_good %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  #multi_boot_standard(col = "answer")
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 3", 
         Condition="Joint")

@@

Language uses symbols expressed in one modality (e.g., the auditory modality, in the case of speech) to communicate about the world, which we perceive through many different sensory modalities. Consider hearing someone yell "bee!" at a picnic, as a honey bee buzzes around the food. Determining reference involves processing the auditory information and linking it with other perceptual signals (e.g., the visual image of the bee, the sound of its wings, the sensation of the bee flying by your arm).

This multimodal integration task takes place in a noisy world. On the auditory side, individual acoustic word tokens are almost always ambiguous with respect to the particular sequence of phonemes they represent, which is due to the inherent variability of how a phonetic category is realized acoustically [@hillenbrand1995]. And some tokens may be distorted additionally by mispronunciation or ambient noise. Perhaps the speaker was yelling "pea" and not "bee". Similarly, a sensory impression may not be enough to make a definitive identification of a visual category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.} Perhaps the insect was a beetle or a fly instead.

How does the brain deal with uncertainty to process and understand language? Previous research showed that the brain encodes speech cues probabilistically, in a way that mirrors the inherent noise/ambiguity in the input. From this probabilistic representation, the brain tends to make the statistically optimal judgement about the identity of the intended categories [e.g., @clayard08; @feldman2009]. Previous work in this line of research has largely focused on testing optimal inference using cues from the auditory modality. A few studies did explore some aspects of audio-visual processing in a probabilistic framework [@bejjanki2011; @kleinschmidt2015]. These studies focused on the case of audio-visual speech perception, that is, recognizing speech categories using cues from the auditory signal accompagnied with lip movements. To our knowledge, however, no study explored whether word recogniton is statistially optimal in a *referential* context, that is, when the brain has to combine cues from the sound and the meaning/referent. Imagine, for example, that someone is uncertain whether they heard "pea" or "bee", does it make them rely more on the referent (e.g., the object being pointed at)? Vice versa, if they are not sure if they saw a bee or a fly, does that make them rely more on the sound? More importantly, when input in both modalities is uncertain to varying degrees, do they weight each modality according to its relative reliability, or do they over-rely on a particular modality?

Outside the realm of language per se, research on multisensory integration suggests that the brain integrates information from different modalities in a statistically optimal way, that is, the integration results in a higher overall precision, and modalities are weighed according to their relative reliability [@ernst02]. However, it is not a priori obvious whether this finding can be extended to word-referent recognition under multimodal uncertainty.  Indeed, there seems to be at least two fundamental differences between these two cases, and both can influence the cue combination strategy.

First, in the case of optimal multisensory integration, modalities are understood to differ only in terms of their reliability. In a referential conext, however, modalities also differ in terms of their roles in the referential process: the auditory input represents the symbol (i.e., the word) whereas the visual input represents the referent. It has been suggested that because of the referential property of speech, it is a privileged signal for humans from infancy [see @vouloumanos2014 for a review]. Thus, it is possible that the brain does not treat the auditory and visual modalities as equivalent sources of information. Instead, there could be a bias for the auditory modality beyond what is expected from informational reliability alone.

Second, research on optimal multisensory integration typically deals with multisensory associations that are supposed to be *redundant*. An example of such integration is when we determine the size of an object using the visual and haptic modalities [@ernst02], or when we determe the spacial location of a stimulus using the visual and the auditory modalities [@alais04]. In the case of language, however, the multimodal input is rather *arbitrary* [@saussure1916].  For instance, there is no logical connection between the sound "bee" and the corresponding insect. Moreover, variation in the way the sound "bee" is pronounced does not correlate perceptually with variation in the shape (or any other visual property) in the category of bees. It turns out the nature of the multi-modal associations (redundant vs. arbitrary) influences the quality of multi-modal processing in both children and adults: the processing and integration is easier in the case of redundant input, than it in the case of arbitrary input [see @robinson2010 for a review]. Thus, there is a priori no obvious reason multimodal cue combination in the case of arbitrary assoiations (such as the case of word-referent) should be optimal in the same way redundant associations are. For instance, the former is more likely to show patterns of sub-optimality.

TRY TO CONTRAST TWO THINGS: WORK ON PERCEPTUAL INTEGRATION WHEN THERE IS CORRELATION AND WORK OF AUDIO-VISUAL ENCODING OF ARBITRARY ASSOCIATIONS, SHOW THAT THE FIRST IS NOT ARBIRARY, AND MOSTY CONTINUOUS (ANS PRODUCE OPTIMALITY), THE SECOND DOES NOT INVOVLE CATEGORIES OR A SEMANTIC CONTEX, WHERE A WORD REFER TO AN OBJECT, INSTEAD OF BEING JUST ARBITRARY (AND PRODUCE HINDRANCE AND SUB-OPTIMALITY). THUS OUR WORK TRIES TO DEAL WITH CASE

@@

In this paper, we studied the mechanism of processing/combination of audiovisual cues in a referential context, under multiple sources of uncertainty. More precisely, we tested if the cue combination was statistically optimal. Besides, we quantified any sub-optimality that might have been induced, among other things, by the suggested privileged status of speech or by the arbitrariness of the referential associations. To test optimality, we characterized uncertainty in each modality with a probability distribution, and we predicted responses by combining these probabilities in an optimal way. To study possible patterns of sub-optimality, we compared the optimal model to a descriptive model where the parameters of the bimodal probability distributions were fitted to human judgements. We tested the model's predictions in three behavioral experiments where we varied the source of uncertainty. In Experiment 1, audio-visual tokens were ambiguous with respect to their category membership. In Experiment 2, we intervened by adding background noise to the auditory modality, and in Experiment 3, we intervened by adding background noise to the visual modality. In all Experiments, participants were quantitatively near-optimal, though precision dropped slightly in the bimodal condition. Moreover, in Experiment 1 where neither of the modalities was perturbed with background noise, participants weighed auditory and visual cues according to the relative reliability predicted by the optimal model. In other words, we found no evidence for a modality bias towards either the auditory or the visual modality. However, In Experiment 2 and 3, participants over-relied on one modality when the other modality was perturbed with additional noise.  Finally, we discuss the broad implications of these results, especially in relation to some developmental findings in early word learning [e.g., @stager1997].

# Model

In this section we describe the multimodal integration model. First, we briefly introduce the experimental paradigm, and second we explain how this paradigm can be modeled in a probabilistic framework.

@@

We introduce a new task that tests word recognition in a referential context.  We use two visual categories (cat and dog) and two auditory categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is set between the auditory and the visual categories, leading to two audio-visual word categories (e.g., dog-/aba/, cat-/ada/).

In each trial, participants are presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test stimulus (Figure\ \@ref(fig:task)). The test stimulus may differ from the target in both the auditory and the visual components.  After these two presentations, participants press "same" or "different".

@@

This itask is similar to the task introduced by @sloutsky2003. However, in that study, as well as subsequent ones, participants were asked whether or not the two audio-visual presentations were *identical*. Such task would allow us to probe audio-visual encoding, but not necessarily speech perception. The latter requires---in addition to the perceptual encoding---a categorical judgement, i.e, determining wheather two similar tokens are members of the same phonological/semantic category. Thus, our task is category-based: participants are asked to press "same" if they think the second item (the test) belonged to the same category as the first (target) (e.g.,  dog-/aba/), even if there is a slight difference in the word, in the referent, or in both. They are instructed to press "different" only if they think that the second stimulus was an instance of the other word category (cat-/ada/).

We assume, for the sake of simplicity, that category tokens vary along a single dimension. Phonetic tokens are sampled from an acoustic continuum linking /aba/ to /aba/, and obtained by continuously varying the value of the second formant [@vroomen2004]. Visual tokens are sampled from a morph continuum linking the picture of a dog and the picture of a cat [@freedman2001].

The task also includes trials where pictures are hidden (audio-only) or where sounds are muted (visual-only). These unimodal trials provide us with participants' categorization functions for the auditory and visual categories and are used as inputs to the optimal model, described below.

@@

The basis of our optimal model is that individual probabilistic representations from each modality should be combined optimally. First we describe how probabilistic representations in each modality are derived/fitted, and second we explain how they are combined to yield optimal predictions.

### Unimodal representations

@@

We assume the probability of membership in each category is normally distributed:

@@

with $\beta_a=\frac{\mu_{A_1}-\mu_{A_2}}{\sigma^2_{A}}$, $\beta_{a0}=\frac{\mu^2_{A_2}-\mu^2_{A_1}}{2\sigma^2_{A}}$, and $1+\epsilon_A=\frac{p(A_1)}{p(A_2)}$ is the proportion of the prior probabilities. If the identity of the auditory categories is randomized, and if $A_1$ is the target, then $\epsilon$ measures a response bias to "same" if $\epsilon > 0$, and a response bias to "different" if $\epsilon < 0$.

@@

### Optimal model

In the bimodal condition, participants see audio-visual tokens. They have to combine information from both modalities and make a categorical decision. How can this task be characterized in a formal way, and what would be the optimal behavior? Work on multisensory integration generally considers the outcome of multiplying the distributions derived from each modality [e.g., @ernst02]. However, this way of combining multi-modal cues is more adequate to examine continuous perceptual dimensions (e.g., size). It is less applicable to categorical dimensions (e.g., phonemes) [see @bejjanki2011; @Bankieris17 for futher detail]

>>>>

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Stanford University"

@@

abstract: |
 Identifying the visual referent of a spoken word -- that a particular insect is referred to by the word ``bee'' -- requires both the ability to process and integrate multimodal input and the ability to reason under uncertainty. How do these tasks interact with one another? We introduce a task that allows us to examine how adults identify words under joint uncertainty in the auditory and visual modalities. We propose an ideal observer model of the task which provides an account of how auditory and visual cues are combined optimally. Model predictions are tested in three experiments where word recognition is made under two kinds of uncertainty: category ambiguity and/or distorting noise. In all cases, the optimal model explains much of the variance in human judgments. In particular, when the signal is not distorted with noise, participants weight the auditory and visual cues optimally, that is, according to the relative reliability of each modality. But when one modality has noise added to it, human perceivers systematically prefer the unperturbed modality to a greater extent than the optimal model does. We discuss the broad implications of the results, especially in relation to early language acuquisition.

@@

header-includes:
   #- \usepackage{bibentry}
   - \usepackage[sortcites=false,sorting=none]{biblatex}

@@

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf

citation_package: biblatex

---

```{r load_packages, include = FALSE}
library(papaja)
library(readr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(boot)
library(dplyr)
library(langcog)
theme_set(theme_bw())
library(broom)
#library("knitcitations")
#cleanbib()
#options("citation_format" = "pandoc")
```

@@

#Data from the 3 experiments 
#############################
exp1 <- read_delim("../Analysis/exp1_data_NoAnonym.txt", delim = " ") %>%
  filter(type == "Task") %>%
  mutate(experiment='Experiment1')

exp2 <- read_delim("../Analysis/exp2_data_NoAnonym.txt", delim = " ") %>%
  filter(type == "Task") %>%
  mutate(experiment='Experiment2')

exp3 <- read_delim("../Analysis/exp3_data_NoAnonym.txt", delim = " ") %>%
  filter(type == "Task") %>%
  mutate(experiment='Experiment3')
```

```{r}
summarise <- dplyr::summarise
rename <- dplyr::rename
```

@@

@@

#Exp1
exp1_noProb <- exp1 %>%
  filter(problem=="No")

@@

#Exp2
exp2_noProb <- exp2 %>%
  filter(problem=="No")

@@

#Exp3
exp3_noProb <- exp3 %>%
  filter(problem=="No")

@@

#Summary
sounds_exp1 <- exp1_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

concepts_exp1 <- exp1_good %>%
  filter(condition == "concept") %>%
  group_by(concept_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Visual") %>%
  rename(distance = concept_dist)

joint_exp1 <- exp1_good %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Joint")

@@

#Summary
sounds_exp2 <- exp2_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 2", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

concepts_exp2 <- exp2_good %>%
  filter(condition == "concept") %>%
  group_by(concept_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 2", 
         Condition="Visual") %>%
  rename(distance = concept_dist)

joint_exp2 <- exp2_good %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 2", 
         Condition="Joint")

@@

#Summary
sounds_exp3 <- exp3_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 3", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

concepts_exp3 <- exp3_good %>%
  filter(condition == "concept") %>%
  group_by(concept_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 3", 
         Condition="Visual") %>%
  rename(distance = concept_dist)

joint_exp3 <- exp3_good %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 3", 
         Condition="Joint")

@@

Language uses symbols expressed in one modality (e.g., the auditory modality, in the case of speech) to communicate about the world, which we perceive through many different sensory modalities. Consider hearing someone yell "bee!" at a picnic, as a honey bee buzzes around the food. Identifying a word involves processing the auditory information as well as other perceptual signals (e.g., the visual image of the bee, the sound of its wings, the sensation of the bee flying by your arm). A word is successfully identified when information from these modalities provide convergent evidence. However, word identification takes place in a noisy world, and the cues received through each modality may not provide a definitive answer. On the auditory side, individual acoustic word tokens are almost always ambiguous with respect to the particular sequence of phonemes they represent, which is due to the inherent variability of how a phonetic category is realized acoustically [@hillenbrand1995]. And some tokens may be distorted additionally by mispronunciation or ambient noise. Perhaps the speaker was yelling "pea" and not "bee". Similarly, a sensory impression may not be enough to make a definitive identification of a visual category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.} Perhaps the insect was a beetle or a fly instead.

How does the mind deal with uncertainty to identify the speaker's intended word? To address this question, many researchers in recent years have adopted the rational approach to cognition [@anderson90], which corresponds to Marr's computational level of analysis. This approach has had a tremendous impact not only on speech related research [@Norris08; @clayard08; @feldman2009; @kleinschmidt2015], but also on many other disciplines in the cognitive sciences [for reviews, see @chater06; @Knill04; @tenenbaum11]. Within this framework, researchers typically consider an "ideal observer" model which provides a characterization of the task/goal and shows what the optimal performance should be under this characterization. When there is uncertainty in the input, the ideal observer performs an optimal probabilistic inference. For example, in order to recognize an ambiguous linguistic input, the model uses all available probabilistic knowledge in order to maximize the accuracy of this recognition. Thus, when the task is well specified, the ideal observer model can be seen as a theoretical upper limit on performance. It is often used, not so much as a realistic model of human performance, as much as a baseline against which human performance can be compared [@Geisler2003]. When there is a deviation from the ideal, this can reveal extra constraints on human cognition, such as limitations on working memory or attentional resources.

In this line of research, @clayard08 simulated auditory uncertainty by manipulating the probability distribution of a cue (VOT) that differentiated similar words (e.g., "beach" and "peach"). They found that humans were sensitive to these probabilistic cues and their judgments closely reflected the optimal predictions. In another work, @feldman2009 studied the perceptual magnet effect, that is, the fact that categories influence the way the stimuli are perceived. They showed that this effect can be explained as the consequence of optimally solving the problem of perception under uncertainty. Both studies explored optimal performance under uncertainty from one modality, i.e., the auditory modality. There is however extensive evidence that information from the visual modality, in particular the speaker's facial features,  influences the perception of auditory information [see @Campbell2008 for a review]. @bejjanki2011 offered a mathematical characterization of how cues from speech and lip movements can be optimally combined and showed that human performance during audio-visual phonemic labeling was consistent (at least at the qualitative level) with the behavior of the ideal observer.

The speaker's facial features are not the only visual information that can be used to disambiguate the auditory input, however. For instance, in a referential context, information about the identity of the semantic referent can be integrated with linguistic information to resolve lexical and syntactic ambiguities in speech [e.g., @Eberhard1995; @Tanenhaus1995; @spivey2002]. To our knowledge, however, no study offered an ideal observer analysis of word identification in such context, that is, when the mind has to combine cues from the sound and the referent. Imagine, for example, that someone is uncertain whether they heard "pea" or "bee", does this uncertainty make them rely more on the referent (e.g., the object being pointed at)? Vice versa, if they are not sure if they saw a bee or a fly, does that make them rely more on the sound? More importantly, when input in both modalities is uncertain to varying degrees, do they weight each modality according to its relative reliability (which is the optimal strategy), or do they over-rely on a particular modality (which is a sub-optimal strategy)?

The visual information provided by the referent is arguably different from the visual information provided by the speaker's facial features (such as lip movements). There seems to be at least two fundamental differences between these two cases, and both can influence the cue combination strategy:

<!-- Outside the realm of language per se, research on multisensory integration suggests that the brain integrates information from different modalities in a statistically optimal way, that is, the integration results in a higher overall precision, and modalities are weighed according to their relative reliability [@ernst02]. However, it is not a priori obvious whether this finding can be extended to word-referent recognition under multimodal uncertainty.  Indeed, there seems to be at least two fundamental differences between these two cases, and both can influence the cue combination strategy. -->

First, in the case of audio-visual speech, modalities are understood to differ only in terms of their reliability. In a referential context, however, modalities also differ in terms of their roles in the referential process: the auditory input represents the symbol (i.e., the word) whereas the visual input represents the referent. It has been suggested that because of the referential property of speech, it is a privileged signal for humans starting in infancy [see @vouloumanos2014 for a review].\footnote{There is a debate over whether speech is privileged for infant and adults for the same reasons. Whereas Waxman and colleagues suggest that speech is privileged for both infants and adults because of its ability to refer, Sloutsky and colleagues suggest that speech might not have a referential status from the start. Rather, they argue tha speech seems to be prefered by infants only because of a low level auditory "overshadowing".} Thus, in a referential context, it is possible that we do not treat the auditory and visual modalities as equivalent sources of information. Instead, there could be a bias for the auditory modality beyond what is expected from informational reliability alone.

<!--An example of such integration is when we determine the size of an object using the visual and haptic modalities [@ernst02], or when we determine the spacial location of a stimulus using the visual and the auditory modalities [@alais04] -->
Second, in the case of audio-visual speech, the associations are expected to be perceptually correlated. The expectation for this correlation is such that when there is a mismatch between auditory and visual input, people still integrate them into a unified (but illusory) percept [@mcgurk1976]. In the case of referential language, however, the multimodal association is by nature *arbitrary* [@saussure1916; @greenberg1957].  For instance, there is no logical/perceptual connection between the sound "bee" and the corresponding insect. Moreover, variation in the way the sound "bee" is pronounced is generally not expected to correlate perceptually with variation in the shape (or any other visual property) in the category of bees. Thus, one could imagine that cue combination in the case of arbitrary audio-visual associations (word-referent) is less automatic, more effortful, and therefore less conducive to optimal integration than it is in the case of perceptually correlated associations (as in audio-visual speech perception).

In the current study, we investigate how people combine cues from the auditory and the visual modality to recognize words in a referential context. In particular, we study how this combination is performed under various degrees of uncertainty in both the auditory and the visual modality.  We perform a rational analysis of the task. First we propose an ideal observer model that performs the combination in an optimal fashion. Second we compare the predictions of the optimal model to human responses. Humans can deviate from the ideal for several reasons. For instance, as mentioned above, a sub-optimality can be induced by the suggested privileged status of speech or by the arbitrariness of the referential association.  In order to study possible patterns of sub-optimality, we compare the optimal model (which provides a normative benchmark) to a descriptive model which is identical to the optimal model, except that the parameters in the former are fitted to human responses. <!--We do not perform the ideal observer analysis at the individual level, but at the population level. There might be some individual variation, but this study addresses the overall performance in the population as a whole.-->

We test the ideal observer model's predictions in three behavioral experiments where we vary the source of uncertainty. In Experiment 1, audio-visual tokens are ambiguous with respect to their category membership only. In Experiment 2, we intervene by adding background noise to the auditory modality, and in Experiment 3, we intervene by adding background noise to the visual modality. In all experiments, participants were quantitatively near-optimal, though overall response precision was slightly lower than expected. Moreover, in Experiment 1 where neither of the modalities was perturbed with background noise, participants weighed auditory and visual cues according to the relative reliability predicted by the optimal model. In other words, we found no evidence for a modality bias towards either the auditory or the visual modality. However, in Experiment 2 and 3, participants over-relied on one modality when the other modality was perturbed with additional noise. <!-- In the discussion, we talk about the broad implications of these results and especially in relation to some developmental findings in early word learning [e.g., @stager1997]. -->

<!--The nature of the multi-modal associations (redundant vs. arbitrary) influences the quality of multi-modal processing in both children and adults: the processing and integration is easier in the case of redundant input, than it in the case of arbitrary input [see @robinson2010 for a review]. Thus, there is a priori no obvious reason multimodal cue combination in the case of arbitrary assoiations (such as the case of word-referent) should be optimal in the same way cue combination of redundant associations is. -->
<!--

@@

# Models and predictions

In this section we describe the multimodal integration model. First, we briefly introduce the experimental paradigm, and second we explain how behavior in this paradigm can be characteized optimally with an ideal observer model.

@@

We introduce a new task that tests word recognition in a referential context.  We use two visual categories (cat and dog) and two auditory categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is set between the auditory and the visual categories, leading to two audio-visual word categories (e.g., dog-/aba/, cat-/ada/). In each trial, participants are presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test stimulus (Figure\ \@ref(fig:task)). The test stimulus may differ from the target in both the auditory and the visual components.  After these two presentations, participants press "same" or "different".

@@

This task is similar to the task introduced by @sloutsky2003, and used in subsequent studies in both children and adults [see @robinson2010 for a review]. However, in this original task, participants are asked whether or not the two audio-visual presentations are *identical*. Such task would allow us to probe audio-visual encoding, but not necessarily language understanding. The latter requires---in addition to the perceptual encoding---a categorical judgement, i.e, determining wheather or not two similar tokens are members of the same phonological/semantic category. Thus, our task is category-based: participants are asked to press "same" if they think the second item (the test) belonged to the same category as the first (target) (e.g.,  dog-/aba/), even if there is a slight difference in the sound, in the referent, or in both. They are instructed to press "different" only if they think that the second stimulus was an instance of the other word category (cat-/ada/). The task also includes trials where pictures are hidden (audio-only) or where sounds are muted (visual-only). These unimodal trials provide us with participants' evaluation of the probabilistic information present in the auditory and visual categories. This unimodal distributions are used as inputs to the optimal combination model, described below.

<!--We assume, for the sake of simplicity, that category tokens vary along a single dimension. Phonetic tokens are sampled from an acoustic continuum linking /aba/ to /aba/, and obtained by continuously varying the value of the second formant [@vroomen2004]. Visual tokens are sampled from a morph continuum linking the picture of a dog and the picture of a cat [@freedman2001]. -->

@@

<!--First we describe how probabilistic representations in each modality are derived, and second we explain how they are combined to yield optimal predictions. -->

<!--
### Unimodal condition

@@

We assume a token's probability of membership in each category to be normally distributed:

@@

with $\beta_a=\frac{\mu_{A_1}-\mu_{A_2}}{\sigma^2_{A}}$, $\beta_{a0}=\frac{\mu^2_{A_2}-\mu^2_{A_1}}{2\sigma^2_{A}}$ and $1+\epsilon_A=\frac{p(A_1)}{p(A_2)}$ is the proportion of the prior probabilities. If the identity of the auditory categories is randomized, and if $A_1$ is the target, then $\epsilon$ measures a response bias to "same" if $\epsilon > 0$, and a response bias to "different" if $\epsilon < 0$.

@@

The ideal observer model has to combine probabilistic information from auditory and the visual modalities and make a categorical decision. How can this task be characterized in a formal way, and what would be the optimal behavior? If the stimuli were continuous (e.g., size, location), cues derived from each modality would be characterized with their sensory noise variance only, and (assuming conditional independence) their combined information would be characterized with a simple linear combination rule where the combined estimate is an average of the indivisual estimates, each wighted according to its relative reliability [@ernst02; @alais04]. However, in our case decison is made over categorical variables (phonemes and concepts), the optimal model should take into account, not only the noise variability arround an individual perceptual estimate, but also its categorical variability (i.e., the probability with which the estimate belongs to a given category).
